---
title: "Daily Electricity Demand in Victoria (2015 – 2022) Analysis"
author: "Kai Ging Yong, Wong Yi Wei (Ethan), Yung Qi Chin"
date: "`r Sys.Date()`"
output: html_document
---

# **Daily Electricity Demand in Victoria (2015 – 2022) Analysis**

## Setup

```{r}
#Libraries
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(TSA)
library(fUnitRoots)
library(lmtest)
library(tseries)
library(forecast)
library(LSTS)
```

```{r}
#Code Setup
#Function 1: Multiple plots with test function

plots<-function(plotdata, plotarg){
  
  #Time Series Plot
  plot(plotdata,
       ylab="Electricity Demand(MWh)",
       xlab="Year",
       type='l',
       main="",
       cex.main=1)
  points(plotdata,x=time(plotdata), pch=as.vector(season(plotdata)))
  grid(nx=NULL, ny=NULL, lty=2, col="gray", lwd=1)
  mtext(paste("Time Series Plot for",plotarg),
        side=3,line=-2.8,outer=TRUE,cex=1)
  
  #Q-Q Plot
  qqnorm(plotdata,
       main="",
       cex.main=1)
  qqline(plotdata,
       col="red",
       lwd=1)
  grid(nx=NULL, ny=NULL, lty=2, col="gray", lwd=1)
  mtext(paste("Normal Q-Q Plot for",plotarg),
        side=3,line=-2.8,outer=TRUE,cex=1)
  
  #ACF & PACF Plot
  par(mfrow=c(1,2))
  acf(plotdata,main="")
  pacf(plotdata,main="")
  mtext(paste("ACF & PACF Plot for",plotarg),
        side=3,line=-2.8,outer=TRUE,cex=1)
  
  #Stationary & Normality Test
  cat(paste(plotarg),"Tests","\n")
  print(adf.test(plotdata))
  print(pp.test(plotdata))
  print(shapiro.test(plotdata))
  
}

#Function 2: Residual Plot
res.plots<-function(plotdata, plotarg, lagmax){
  
  #Time Series Plot
  plot(plotdata,xlab='Time',ylab='Residuals',main="",type="l")
  grid(nx=NULL, ny=NULL, lty=2, col="gray", lwd=1)
  points(plotdata,x=time(plotdata), pch=as.vector(season(plotdata)))
  mtext(paste("Time Series Plot for",plotarg),
        side=3,line=-2.8,outer=TRUE,cex=1)
  
  #ACF & PACF Plot
  par(mfrow=c(1,2))
  acf(plotdata,main="",lag.max=lagmax)
  pacf(plotdata,main="",lag.max=lagmax)
  mtext(paste("ACF & PACF Plot for",plotarg),
        side=3,line=-2.8,outer=TRUE,cex=1)
  
}

#Function 3: Residual Analysis
residual.analysis <- function(model, std = TRUE,start = 2, class = c("ARIMA","GARCH","ARMA-GARCH", "fGARCH")[1]){
  
  if (class == "ARIMA"){
    if (std == TRUE){
      res.model = rstandard(model)
    }else{
      res.model = residuals(model)
    }
  }else if (class == "GARCH"){
    res.model = model$residuals[start:model$n.used]
  }else if (class == "ARMA-GARCH"){
    res.model = model@fit$residuals
  }else if (class == "fGARCH"){
    res.model = model@residuals
  }else {
    stop("The argument 'class' must be either 'ARIMA' or 'GARCH' ")
  }
  par(mfrow=c(3,2))
  plot(res.model,type='o',ylab='Standardised residuals', main="Time series plot of standardised residuals")
  abline(h=0)
  hist(res.model,main="Histogram of standardised residuals")
  qqnorm(res.model,main="QQ plot of standardised residuals")
  qqline(res.model, col = 2)
  acf(res.model,main="ACF of standardised residuals")
  print(shapiro.test(res.model))
  k=0
  par(mfrow=c(1,1))
  Box.Ljung.Test(res.model, lag = 30)
}

#Function 4: AIC & BIC Score sort function
sort.score <- function(x, score = c("bic", "aic")){
  if (score == "aic"){
    x[with(x, order(AIC)),]
  } else if (score == "bic") {
    x[with(x, order(BIC)),]
  } else {
    warning('score = "x" only accepts valid arguments ("aic","bic")')
  }
}
```

------------------------------------------------------------------------

## 1.0 Import Data and Preprocessing

```{r}
#Import data
electricity<-read.csv("complete_dataset.csv")

#Check data class
class(electricity) 

#Check data structure
str(electricity)

#Random 5 observations
electricity[sample(nrow(electricity),5), ]
```

```{r}
#Create year & month columns
electricity$date <- as.character(electricity$date)
electricity$year <- as.integer(substr(electricity$date, 1, 4))
electricity$month <- as.integer(substr(electricity$date, 6, 7))

#Generate new dataframe
electricity <- electricity %>%
  group_by(year, month) %>%
  summarize(total_demand = sum(demand))
electricity <- electricity[-nrow(electricity), ]

#Random 5 observations
electricity[sample(nrow(electricity),5), ]

#Summary of dataframe
summary(electricity)
```

```{r}
#Convert to time series object
electricity.ts = ts(electricity$total_demand,start=c(2015,1),frequency=12)
```

------------------------------------------------------------------------

## 2.0 Data Visualization and Descriptive Analysis

```{r}
plots(electricity.ts,"Electricity Demand Series")
```

```{r}
#Scatter plot
plot(y=electricity.ts,
     x=zlag(electricity.ts),
     xlab="Electricity Demand (MWH)",
     ylab="Previous Month Electricty Demand (MWh)",
     main="Scatter Plot for Electricity Demand Series")

abline(lm(electricity.ts~zlag(electricity.ts), data=electricity.ts),col="red")

grid(nx=NULL, ny=NULL, lty=2, col="gray", lwd=1)
```

```{r}
#Correlation
y=electricity.ts
x=zlag(electricity.ts)
index=2:length(x)
cat("Correlation:",cor(y[index],x[index]))
```

------------------------------------------------------------------------

## 3.0 Data Preparation & Modelling

### 3.1 Box Cox Transformation

```{r}
#Check for negative values
cat("Sum of Negative Values:", sum(electricity.ts<0),"\n\n")
cat("Data Summary:","\n")
summary(electricity.ts)
```

```{r}
#Box Cox transformation
electricity.bc <- BoxCox.ar(electricity.ts)

mtext(paste("Figure 5: Box-Cox Plot for Electricity Demand Series"),
      side=3,
      line=-2.8,
      outer=TRUE,
      cex=1)
```

```{r}
#Lambda & CI values
lambda <- electricity.bc$lambda[which(max(electricity.bc$loglike) == electricity.bc$loglike)]
cat("CI:",electricity.bc$ci,"\n")
cat("Lambda:",lambda)

BC.electricity = ((electricity.ts^lambda)-1)/lambda
```

```{r}
#BC Time Series plot
plots(BC.electricity, "Box Cox Transformed Electricity Demand Series")
```

### 3.2 SARIMA Approach

```{r}
#(0, 0, 0)
model1 = Arima(electricity.ts,order=c(0,0,0),seasonal=list(order=c(0,1,0), period=12))
res.model1 = residuals(model1)
res.plots(res.model1,"Model 1 Residuals",48)
```

```{r}
#(0, 1, 1) BC
model2 = Arima(electricity.ts,order=c(0,0,0),seasonal=list(order=c(0,1,1), period=12))
res.model2 = residuals(model2)
res.plots(res.model2, "Model 2 Residuals", 36)
```

```{r}
#(0, 1, 1) TS
model3 = Arima(BC.electricity,order=c(0,0,0),seasonal=list(order=c(0,1,1), period=12))
res.model3 = residuals(model3)
res.plots(res.model3, "Model 3 Residuals", 36)
```

```{r}
#(5,0,4) BC
model4 = Arima(BC.electricity,order=c(5,0,4),seasonal=list(order=c(0,1,1), period=12))
res.model4 = residuals(model4)
res.plots(res.model4, "Model 4 Residuals" , 36)
```

------------------------------------------------------------------------

## 4.0 Model Specification

### 4.1 Extended Correlation Function (EACF)

```{r}
#EACF
cat("EACF Table","\n")
eacf(res.model3)
```

### 4.2 Subset ARMA Model

```{r}
#BIC
par(mfrow=c(1,1))
bic_table = armasubsets(y=res.model3,nar=10,nma=10,y.name='p',ar.method='ols')
plot(bic_table)
mtext("BIC Table",side=3,line=-1.5,outer=TRUE,cex=1)
```

------------------------------------------------------------------------

## 5.0 Parameter Estimation

### 5.1 SARIMA (4,0,5)

```{r}
#ML
model5_405 = Arima(BC.electricity,order=c(4,0,5),seasonal=list(order=c(0,1,1), period=12),method = "ML")

coeftest(model5_405)

residual.analysis(model = model5_405)
```

```{r}
#CSS
model5_405_CSS = Arima(BC.electricity,order=c(4,0,5),seasonal=list(order=c(0,1,1), period=12),method = "CSS")

coeftest(model5_405_CSS)

residual.analysis(model = model5_405_CSS)
```

```{r}
#CSS-ML
model5_405_CSSML = Arima(BC.electricity,order=c(4,0,5),seasonal=list(order=c(0,1,1), period=12),method = "CSS-ML")

coeftest(model5_405_CSSML)

residual.analysis(model = model5_405_CSSML)
```

### 5.2 SARIMA (0,0,2)

```{r}
#ML
model5_002 = Arima(BC.electricity,order=c(0,0,2),seasonal=list(order=c(0,1,1), period=12),method = "ML")

coeftest(model5_002)

residual.analysis(model = model5_002)
```

```{r}
#CSS
model5_002_CSS = Arima(BC.electricity,order=c(0,0,2),seasonal=list(order=c(0,1,1), period=12),method = "CSS")

coeftest(model5_002_CSS)

residual.analysis(model = model5_002_CSS)
```

```{r}
#CSS-ML
model5_002_CSSML = Arima(BC.electricity,order=c(0,0,2),seasonal=list(order=c(0,1,1), period=12),method = "CSS-ML")

coeftest(model5_002_CSSML)

residual.analysis(model = model5_002_CSSML)
```

### 5.3 SARIMA (0,0,3)

```{r}
#ML
model5_003 = Arima(BC.electricity,order=c(0,0,3),seasonal=list(order=c(0,1,1), period=12),method = "ML")

coeftest(model5_003)

residual.analysis(model = model5_003)
```

```{r}
#CSS
model5_003_CSS = Arima(BC.electricity,order=c(0,0,3),seasonal=list(order=c(0,1,1), period=12),method = "CSS")

coeftest(model5_003_CSS)

residual.analysis(model = model5_003_CSS)
```

```{r}
#CSS-ML
model5_003_CSSML = Arima(BC.electricity,order=c(0,0,3),seasonal=list(order=c(0,1,1), period=12),method = "CSS-ML")

coeftest(model5_003_CSSML)

residual.analysis(model = model5_003_CSSML)
```

### 5.4 SARIMA (1,0,2)

```{r}
#ML
model5_102 = Arima(BC.electricity,order=c(1,0,2),seasonal=list(order=c(0,1,1), period=12),method = "ML")

coeftest(model5_102)

residual.analysis(model = model5_102)
```

```{r}
#CSS
model5_102_CSS = Arima(BC.electricity,order=c(1,0,2),seasonal=list(order=c(0,1,1), period=12),method = "CSS")

coeftest(model5_102_CSS)

residual.analysis(model = model5_102_CSS)
```

```{r}
#CSS-ML
model5_102_CSSML = Arima(BC.electricity,order=c(1,0,2),seasonal=list(order=c(0,1,1), period=12),method = "CSS-ML")

coeftest(model5_102_CSSML)

residual.analysis(model = model5_102_CSSML)
```

### 5.5 SARIMA (1,0,3)

```{r}
#ML
model5_103 = Arima(BC.electricity,order=c(1,0,3),seasonal=list(order=c(0,1,1), period=12),method = "ML")

coeftest(model5_103)

residual.analysis(model = model5_103)
```

```{r}
#CSS
model5_103_CSS = Arima(BC.electricity,order=c(1,0,3),seasonal=list(order=c(0,1,1), period=12),method = "CSS")

coeftest(model5_103_CSS)

residual.analysis(model = model5_103_CSS)
```

### 5.6 SARIMA (8,0,6)

```{r}
#ML
model5_806 = Arima(BC.electricity,order=c(8,0,6),seasonal=list(order=c(0,1,1), period=12),method = "ML")

coeftest(model5_806)

residual.analysis(model = model5_806)
```

```{r}
#CSS
model5_806_CSS = Arima(BC.electricity,order=c(8,0,6),seasonal=list(order=c(0,1,1), period=12),method = "CSS")

coeftest(model5_806_CSS)

residual.analysis(model = model5_806_CSS)
```

```{r}
#CSS-ML
model5_806_CSSML = Arima(BC.electricity,order=c(8,0,6),seasonal=list(order=c(0,1,1), period=12),method = "CSS-ML")

coeftest(model5_806_CSSML)

residual.analysis(model = model5_806_CSSML)
```

### 5.7 SARIMA (2,0,6)

```{r}
#ML
model5_206 = Arima(BC.electricity,order=c(2,0,6),seasonal=list(order=c(0,1,1), period=12),method = "ML")

coeftest(model5_206)

residual.analysis(model = model5_206)
```

```{r}
#CSS
model5_206_CSS = Arima(BC.electricity,order=c(2,0,6),seasonal=list(order=c(0,1,1), period=12),method = "CSS")

coeftest(model5_206_CSS)

residual.analysis(model = model5_206_CSS)
```

```{r}
#CSS-ML
model5_206_CSSML = Arima(BC.electricity,order=c(2,0,6),seasonal=list(order=c(0,1,1), period=12),method = "CSS-ML")

coeftest(model5_206_CSSML)

residual.analysis(model = model5_206_CSSML)
```

### 5.8 SARIMA (10,0,6)

```{r}
#ML
model5_1006 = Arima(BC.electricity,order=c(10,0,6),seasonal=list(order=c(0,1,1), period=12),method = "ML")

coeftest(model5_1006)

residual.analysis(model = model5_1006)
```

```{r}
#CSS
model5_1006_CSS = Arima(BC.electricity,order=c(10,0,6),seasonal=list(order=c(0,1,1), period=12),method = "CSS")

coeftest(model5_1006_CSS)

residual.analysis(model = model5_1006_CSS)
```

```{r}
#CSS-ML
model5_1006_CSSML = Arima(BC.electricity,order=c(10,0,6),seasonal=list(order=c(0,1,1), period=12),method = "CSS-ML")

coeftest(model5_1006_CSSML)

residual.analysis(model = model5_1006_CSSML)
```

------------------------------------------------------------------------

## 6.0 Goodness-of-Fit Metrics

### 6.1 Akaike Information Criterion (AIC) Scores

```{r}
sort.score(AIC(model5_405,
               model5_002,
               model5_003,
               model5_102,
               model5_103,
               model5_206), score="aic")
```

### 6.2 Bayesian Information Criterion (BIC) Scores

```{r}
sort.score(BIC(model5_405,
               model5_002,
               model5_003,
               model5_102,
               model5_103,
               model5_206),score="bic")
```

### 6.4 Accuracy Measures

```{r}
model5_504_a<-accuracy(model5_405)[1:7]
model5_002_a<-accuracy(model5_002)[1:7]
model5_003_a<-accuracy(model5_003)[1:7]
model5_102_a<-accuracy(model5_102)[1:7]
model5_103_a<-accuracy(model5_103)[1:7]
model5_206_a<-accuracy(model5_206)[1:7]

models_a<-data.frame(rbind(model5_504_a,
                           model5_002_a,
                           model5_003_a,
                           model5_102_a,
                           model5_103_a,
                           model5_206_a))

colnames(models_a) <- c("ME", "RMSE", "MAE", "MPE", "MAPE", "MASE", "ACF1")

rownames(models_a) <- c("SARIMA(4,0,5)x(0,1,1)_12", 
                        "SARIMA(0,0,2)x(0,1,1)_12", 
                        "SARIMA(0,0,3)x(0,1,1)_12", 
                        "SARIMA(1,0,2)x(0,1,1)_12", 
                        "SARIMA(1,0,3)x(0,1,1)_12",
                        "SARIMA(2,0,6)x(0,1,1)_12")

models_a<-round(models_a, digits = 3)

models_a %>% arrange(MASE)
```

### 6.5 Over-parameterized model: SARIMA(2,0,3)x(0,1,1)\_12

```{r}
#ML
model5_203 = Arima(BC.electricity,order=c(2,0,3),seasonal=list(order=c(0,1,1), period=12),method = "ML")

coeftest(model5_203)

residual.analysis(model = model5_203)
```

```{r}
#CSS
model5_203_CSS = Arima(BC.electricity,order=c(2,0,3),seasonal=list(order=c(0,1,1), period=12),method = "CSS")

coeftest(model5_203_CSS)

residual.analysis(model = model5_203_CSS)
```

### 6.5 Over-parameterized model: SARIMA(1,0,4)x(0,1,1)\_12

```{r}
#ML
model5_104 = Arima(BC.electricity,order=c(1,0,4),seasonal=list(order=c(0,1,1), period=12),method = "ML")

coeftest(model5_104)

residual.analysis(model = model5_104)
```

```{r}
#CSS
model5_104_CSS = Arima(BC.electricity,order=c(1,0,4),seasonal=list(order=c(0,1,1), period=12),method = "CSS")

coeftest(model5_104_CSS)

residual.analysis(model = model5_104_CSS)
```

### 6.6  AIC and BIC Scores (With Overparameterized Models)

```{r}
sort.score(AIC(model5_405,
               model5_002,
               model5_003,
               model5_102,
               model5_103,
               model5_206,
               model5_203,
               model5_104), score="aic")
```

```{r}
sort.score(BIC(model5_405,
               model5_002,
               model5_003,
               model5_102,
               model5_103,
               model5_206,
               model5_203,
               model5_104), score="bic")
```

## 7.0 Forecasting

```{r}
model5_103.electricity = Arima(electricity.ts,order=c(1,0,3),seasonal=list(order=c(0,1,1), period=12), lambda = 2, method = "ML")

model5_103_frc = forecast(model5_103.electricity, lambda = 2, h = 10)
model5_103_frc
plot(model5_103_frc, 
     main="Next 10 Months Forecast for Electricity Demand Series from SARIMA(1,0,3)(0,1,1)", cex.main=0.9)
```
